---
title: "HSDSArray -- DelayedArray backend for Remote HDF5"
author: "Samuela Pollack"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{HSDSArray DelayedArray backend}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

```{r setup,echo=FALSE,results="hide"}
suppressPackageStartupMessages({
suppressMessages({
library(rhdf5client2)
library(DelayedArray)
})
})
```
# Using the DelayedArray infrastructure

A remote dataset is accessed by giving the URL of the server, the
type of the server (either `hsds` or `h5serv`), the file domain (path 
to the HDF5 file) and the full path to the dataset inside the HDF5 file. 

## Interface to h5serv (Channing Remote Data Server)

```{r}
da <- HSDSArray('http://h5s.channingremotedata.org:5000', 'h5serv', 
      'tenx_full.h5s.channingremotedata.org', '/newassay001')
da
```

Here we have defined the R image of the data to be the
transpose of the image in HDF5.  So neurons are columns.

```{r}
apply(da[,1:4],2,sum)
```

## Interface to HSDS (HDF Object Store)
```{r lkdela3}
da <- HSDSArray('http://hsdshdflab.hdfgroup.org', 'hsds', 
      '/shared/bioconductor/tenx_full.h5', '/newassay001')
da
```

Again we have DelayedArray capabilities.
```{r lkdel3}
apply(da[,1:4],2,sum)
```
